Katherine Analysis

### 2018-03-15 Katherine Analysis.Rmd

AICres <- model.sel(mod1,mod2,mod3)
r.squaredLR(mod1)
summary(mod1)
confint(mod1)
```

@@

### Linear discriminate analysis
```{r}
# Discriminant function analysis is used (general LDA format) because i have a cat. dependent and a binary independent. (it needs binary or continuous!)
# you might have to bin covers because it only works with categorical, so 0-5% none, 6-40 low etc.
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("None", "Med_Low", "Med_High", "Overtaken"))
table(FloristicSurvey$GM_Coverage_category)
# [0-20] None
# ]20-50] Medium low
# ]50-80] Medium high
# ]80-100] Overtaken
```

@@

# Plot results
ggplot(data=FloristicSurveyLDAval,aes(x=LD1,y=LD2,group=FloristicSurveysmall$GM_Coverage_category))+
  stat_ellipse(geom="polygon",aes(colour=FloristicSurveysmall$GM_Coverage_category),fill=NA,size=1.2,alpha=0.3)+
  stat_ellipse(geom="polygon",aes(fill=FloristicSurveysmall$GM_Coverage_category,colour=FloristicSurveysmall$GM_Coverage_category),size=1.2,alpha=0.3)+
  geom_point(aes(shape=FloristicSurveysmall$GM_Coverage_category,fill=FloristicSurveysmall$GM_Coverage_category,colour=FloristicSurveysmall$GM_Coverage_category),size=I(4),alpha=I(0.8))+
  xlab("LD Axis 1")+ylab("LD Axis 2")+theme_simple() +
  theme(legend.title=element_blank())

@@

par(mfrow=c(3,2))
plot(fit,cex=0.4,main="Clustering with populations")
plot(cut(fitd, h=10)$upper, main="Upper tree of cut at h=10")
plot(cut(fitd, h=10)$lower[[1]], main="First branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[2]], main="Second branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[3]], main="Third branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[4]], main="Fourth branch of lower tree with cut at h=10")
```

### NMDS 
```{r}
# https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/
# https://oliviarata.wordpress.com/2014/04/17/ordinations-in-ggplot2/
#Make a matrix with no row or column equal to 0 (do not enclude the env variable (GM COVERAGE))
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("Low", "Med_Low", "Med_High", "High"))
table(FloristicSurvey$GM_Coverage_category)
M <- as.matrix(FloristicSurvey[1:198,4:56])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
M<-M[-c(130,133),]
M<-M[,-45]
GM_coverage_df <- data.frame(M[,52])
M<-M[,-52]
class(M) <- "numeric"

# with vegdist from Bray to seroeson: add binary = T 
dist_FloristicSurvey <- vegdist(M, method = "bray", binary = T)
#The metaMDS analysis could have done the distance matrix internally but i would rather control it since i have presence/abscence
meta.nmds.FloristicSurvey2D <- metaMDS(dist_FloristicSurvey, k=2, trymax = 1000)

@@

## finally plotting.
mult <- 1 #multiplier for the arrows and text for envfit below. You can change this and then rerun the plot command.
ggplot(data = FloristicSurvey.NMDS.data, aes(y = NMDS2, x = NMDS1))+ 
    geom_path(data = df_ell.FSurvey.GM_coverage, aes(x = NMDS1, y = NMDS2, group = df_ell.FSurvey.GM_coverage$GM_coverage, color=df_ell.FSurvey.GM_coverage$GM_coverage))+ 
    geom_point( aes(color = FloristicSurvey.NMDS.data$GM_Coverage), size = 1)+ 
    scale_color_manual(values = c(14,5,19,2))+ 
    coord_cartesian(xlim = c(-1,1.5))+
  theme_simple()+ guides(color=guide_legend("Garlic Mustard Coverage"))
```

### PCA
```{r}
FloristicSurvey<-FloristicSurvey[,-56]
FloristicSurvey.pca <- prcomp(FloristicSurvey[,4:(ncol(FloristicSurvey))],center = TRUE) 
summary(FloristicSurvey.pca)
fviz_eig(FloristicSurvey.pca,addlabels=TRUE,hjust = -0.3)+theme_simple()

round(FloristicSurvey.pca$sdev^2/sum(FloristicSurvey.pca$sdev^2)*100,2)
# About 25% explained by my first 2 eigenvectors. 50% by the first 6.
```
### Modeling (redundancy analysis)
species comp ~ GMcov * pop + soil characteristics inside and outside as covariates? 
```{r}

@@

var$Population <- as.factor(var$Population)

>>>>

AICres <- model.sel(mod1,mod2,mod3)
r.squaredLR(mod1)
summary(mod1)
confint(mod1)

@@

### Linear discriminate analysis
```{r}
# Discriminant function analysis is used (general LDA format) because i have a cat. dependent and a binary independent. (it needs binary or continuous!)
# you might have to bin covers because it only works with categorical, so 0-5% none, 6-40 low etc.
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("None to Low","Medium", "High", "Overtaken"))
table(FloristicSurvey$GM_Coverage_category)
# [0-20] None
# ]20-40] low
# ]40-60] Medium 
# ]60-80] high
# ]80-100] Overtaken
```

@@

# Plot results
LDA_of_Floristic_Composition<-ggplot(data=FloristicSurveyLDAval,aes(x=LD1,y=LD2,group=FloristicSurveysmall$GM_Coverage_category))+
  stat_ellipse(geom="polygon",aes(colour=FloristicSurveysmall$GM_Coverage_category),fill=NA,size=1.2,alpha=0.3)+
  stat_ellipse(geom="polygon",aes(fill=FloristicSurveysmall$GM_Coverage_category,colour=FloristicSurveysmall$GM_Coverage_category),size=1.2,alpha=0.3)+
  geom_point(aes(shape=FloristicSurveysmall$GM_Coverage_category,fill=FloristicSurveysmall$GM_Coverage_category,colour=FloristicSurveysmall$GM_Coverage_category),size=I(4),alpha=I(0.8))+
  xlab("LD Axis 1")+ylab("LD Axis 2")+
  ggtitle("LDA of Floristic Composition \n in a Garlic Mustard Population")+
  theme_bw()+
  theme(legend.title=element_blank(),plot.title = element_text(hjust = 0.5,face="bold", size=20), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(colour = "black"),axis.title.x = element_text(size=18),axis.title.y = element_text(size=18),axis.text.y = element_text(size=10),panel.border = element_rect(colour = "black"),strip.background = element_rect(colour = "black",fill="white"),strip.text.x = element_text(size = 12))+scale_color_manual(values=c("green3"," blue3","orange3","red3"))+ scale_fill_manual(values=c("green3","blue3","orange3","red3"))
ggsave(file="LDA_of_Floristic_Composition.epx", plot=LDA_of_Floristic_Composition, width=10, height=8)

@@

par(mfrow=c(3,2))
plot(fit,cex=0.4,main="Clustering with coverage")
plot(cut(fitd, h=10)$upper, main="Upper tree of cut at h=10")
plot(cut(fitd, h=10)$lower[[1]], main="First branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[2]], main="Second branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[3]], main="Third branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[4]], main="Fourth branch of lower tree with cut at h=10")
```

### NMDS 
```{r}
# https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/
# https://oliviarata.wordpress.com/2014/04/17/ordinations-in-ggplot2/
#Make a matrix with no row or column equal to 0 (do not enclude the env variable (GM COVERAGE))
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("None to Low","Medium", "High", "Overtaken"))
table(FloristicSurvey$GM_Coverage_category)
M <- as.matrix(FloristicSurvey[1:198,4:56])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
M<-M[-c(130,133),]
M<-M[,-45]
GM_coverage_df <- data.frame(M[,52])
M<-M[,-52]
class(M) <- "numeric"

# with vegdist from Bray to seroeson: add binary = T 
dist_FloristicSurvey <- vegdist(M, method = "bray", binary = T)
meta.nmds.FloristicSurvey2D <- metaMDS(dist_FloristicSurvey, k=2, trymax = 1000)

@@

## finally plotting.
mult <- 1 #multiplier for the arrows and text for envfit below. You can change this and then rerun the plot command.
ggplot(data = FloristicSurvey.NMDS.data, aes(y = NMDS2, x = NMDS1))+ 
    geom_path(data = df_ell.FSurvey.GM_coverage, aes(x = NMDS1, y = NMDS2, group = df_ell.FSurvey.GM_coverage$GM_coverage, color=df_ell.FSurvey.GM_coverage$GM_coverage))+ 
    geom_point( aes(color = FloristicSurvey.NMDS.data$GM_Coverage), size = 1)+ 
    scale_color_manual(values = c(14,5,19,2))+ 
    coord_cartesian(xlim = c(-1,1.5))+
  theme_simple()+ guides(color=guide_legend("Garlic Mustard Coverage"))+
  ggtitle("LDA of Floristic Composition \n in a Garlic Mustard Population")+
  theme_bw()+
  theme(legend.title=element_blank(),plot.title = element_text(hjust = 0.5,face="bold", size=20), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank(),axis.line = element_line(colour = "black"),axis.title.x = element_text(size=18),axis.title.y = element_text(size=18),axis.text.y = element_text(size=10),panel.border = element_rect(colour = "black"),strip.background = element_rect(colour = "black",fill="white"),strip.text.x = element_text(size = 12))+scale_color_manual(values=c("green3"," blue3","orange3","red3"))+ scale_fill_manual(values=c("green3","blue3","orange3","red3"))
ggsave(file="LDA_of_Floristic_Composition.epx", plot=LDA_of_Floristic_Composition, width=10, height=8)
```

### PCA
```{r}
FloristicSurvey.pca <- prcomp(FloristicSurvey[,4:(ncol(FloristicSurvey)-1)],center = TRUE) 
summary(FloristicSurvey.pca)
factoextra::fviz_eig(FloristicSurvey.pca,addlabels=TRUE,hjust = -0.3)+theme_simple()
library(ggfortify)
autoplot(FloristicSurvey.pca, data = FloristicSurvey, colour = "GM_Coverage_category", label.size = 3)+theme_simple()
round(FloristicSurvey.pca$sdev^2/sum(FloristicSurvey.pca$sdev^2)*100,2)
# About 25% explained by my first 2 eigenvectors. 50% by the first 6.
```

### Modeling (redundancy analysis)
species comp ~ GMcov * pop + soil characteristics inside and outside as covariates? 
```{r}

@@

var$Population <- as.factor(var$Population)
var$GM_Coverage<-as.numeric(var$GM_Coverage)
sp.comp<-as.matrix(sp.comp)
